# Data Visualization


## Data Visualization in Base R

Basic R plots in base R such as **hist()** and **plot()**

```{r}
library(tidyverse)
```


```{r}
#We are using the same MN_Data as earlier...
MN_Data <- read_csv("MN_county_tract_pooled.csv")
attach(MN_Data) #by attaching a data frame, we can now refer to the variables in the data frame by their names alone, rather than as components of the data frame (eg. MN_Data$median_rent2016)

hist(median_rent2016, main="Histogram of Median Rent in 2016 for MN Census tracts", xlab="median rent in 2016")
plot(popdensity2010, nonwhite_share2010,
     main="Median rent in 2016 for MN Census tracts",
        xlab="population density in 2010 for MN Census tracts",
        ylab="non white share population 2010 for MN Census tracts") # when you put the name of a two-variable data frame, it will plot the first column as X and second column as Y 
#xlab and ylab add x and y labels. Main gives the graph a title

```

## ggplot2 basics

the package ggplot has lots of extra plot code to create unique fun plots

```{r}
library(ggplot2) 
```

ggplot2 is based on the **grammar of graphics**, it is the idea that you can build graphs from 1) a data set, 2) a set of geoms, 3) and a coordinate system. 

geoms are visual marks that represent data points. Geoms tell the plot how you want to display the data. Examples are geom_histogram, geom_area, geom_bar etc...


To display data values, map variables in the data set to **aesthetic** properties of the geom like **size, color, and x and y locations**.


Basic format example: ggplot(data=mpg, aes(x=cty, y=hwy))
"Geoms" represent data points, the geom's aesthetic properties represent variables. 

## Labels

Labels and titles can be added to graphs using **ggtitle()**, **xlab()**, and **ylab()**...or **labs()** for all three.

```{r}

t <- ggplot(mpg, aes(hwy))

t + ggtitle("New Plot Title") #adds a main title above the plot

t + xlab("New X Label") #Change the label on the X axis

t + ylab("New Y Label") #Change the label on the Y axis

t + labs(title="New title", x="New X", y="New Y") #all of the above labels
```

## Theme Options

```{r}
r <- ggplot(mpg, aes(cty,hwy))
r + geom_blank() #Blank graph
r + theme_bw() #creates a black and white background theme
r + theme_grey() #creates a grey background theme
r + theme_classic() #creates a white background theme
r + theme_minimal() #creates a minimal white and grey background theme
```

## One Variable Continuous 

```{r}
a <- ggplot(mpg, aes(hwy))
a + geom_area(stat = "bin",color="blue",fill="pink") 
 # [options: x, y, alpha, color, fill, size] #used fill option above creating the pink area, color added the blue outline
a + geom_dotplot(color="yellow") 
 # [options: x,y,alpha, color, fill]
a + geom_histogram(binwidth=5, color="white")+ #color="white" helps bars not run together
  labs(title="mpg histogram", x="mpg")+ #adds title and x labels
  theme_minimal() #gets rid of grey background
# [options: x,y,alpha, color,fill,linetype,size, weight] 
```

## One Variable Discrete

Using the mpg data set previously used in the ggplot2 package, the variable fl is 'fuel type' from fuel economy data from 1999 to 2008 for 38 popular models of cars. 

```{r}
b <- ggplot(mpg,aes(fl))
b + geom_bar(fill="pink")+ #bars become pink
  coord_flip()+ #flips bars horizontal
  ggtitle("mpg, fuel type")+ #adds title
  xlab("fuel type")+
  theme_bw() #adds a black and white background theme to graph
  #[options: x,alpha, color, fill, linetype, size, weight] #used fill option in above graph.

```
## Two Variable Continuous X, Continuous Y

**geom_point()** plots a scatterplot the points in the graph as they are without any variation

**geom_jitter()** adds a small amount of random variation to the location of each point, and is a useful way of handling overplotting caused by discreteness in smaller datasets.

**geom_smooth()** helps point out a pattern with a smooth line

link to shape -> number codes used below: http://sape.inf.usi.ch/quick-reference/ggplot2/shape


link to some unique color options: http://sape.inf.usi.ch/quick-reference/ggplot2/colour
```{r}
f <- ggplot(mpg, aes(cty,hwy))
f + geom_blank() #Blank graph
f + geom_jitter(
  color="orange",
  shape=13 #unique bullsey shape
)+
  xlab("city miles per gallon")+
  ylab("highway miles per gallon")
 # [x,y,alpha, color, fill, shape, size]


f + geom_point(
  color="purple",#outline is purple
        fill="#345698", #numbers correspond to a color
        shape=22, #shape is a square
        alpha=0.5, #transparency is 50%, helps show overlap in this example
        size=6)+
  xlab("city miles per gallon")+
  ylab("highway miles per gallon") #Size of shape

f + geom_smooth(model = lm, color="darkseagreen2") #found color using link above
 # [x,y,alpha, color, fill, ,linetype, size, weight] #other customization options
```

```{r}
ggplot(mpg, aes(hwy, cty)) # This creates a basic graph that you can add layers to to finish it. 

ggplot(mpg, aes(hwy, cty)) +
  geom_point(aes(color = cyl)) +
  geom_smooth(method="lm") +
  coord_cartesian() +
  scale_color_gradient() +
  theme_bw()+
  xlab("city miles per gallon")+
  ylab("highway miles per gallon")
```

## Two Variable Discrete X, Continuous Y


```{r}
g <- ggplot(mpg,aes(class,hwy))
g + geom_boxplot(color="coral")+
  ylab("highway miles per gallon")+
  xlab("type of car")
  #[Options for inside geom_bar parenthesis: lower, middle, upper, x, ymax, ymin, alpha, color, fill, linetype, shape, size, weight] #used color
```

## Two Variable Discrete X, Discrete Y

The elect data set used below is data from the 2016 presidential election. The data set, prepared by Alicia Johnson (Macalester College), combines 2008/2012/2016 county-level election returns from [Tony McGovern  on github](https://github.com/tonmcg/County_Level_Election_Results_12-16), county-level demographics from the `df_county_demographics` data set within the `choroplethr` R package, and red/purple/blue state designations from [http://www.270towin.com/](http://www.270towin.com/).

The `winrep_2016` variable indicates whether or not the Republican (Trump) won the county in 2016'

```{r}
elect <- read_csv("https://www.macalester.edu/~ajohns24/data/electionDemographics16.csv") #reads in elect data

#stacked bar plot

ggplot(elect, aes(x=winrep_2016, fill=StateColor)) + 
  geom_bar()+
  xlab("Did Trump win the county in 2016?")

#side-by-side bar plot
ggplot(elect, aes(x=winrep_2016, fill=StateColor)) + 
  geom_bar(position="dodge")+
  xlab("Did Trump win the county in 2016?")
```
  
## Two Variable Continuous Function

The next data set used is the **economics** data set in the *ggplot2* package. This is from the US economics time series data. 

```{r}

j <- ggplot(economics, aes(date, unemploy)) #date is the month of data collection, unemploy is the number of unemployed in thousands.
j + geom_area()+
  xlab("month of data collection")+
  ylab("number of unemployed (in thousands)")
j + geom_line()+
  xlab("month of data collection")+
  ylab("number of unemployed (in thousands)")
```

## Maps

The next data set is from base R. It is called USArrests. It contains statistics in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973.

```{r}
#install *maps* package
library(maps)
data <- data.frame(murder = USArrests$Murder,state=tolower(rownames(USArrests))) #the data is the Murder: number of murder arrests per 100,000. state is the 
map <- map_data("state") #map by state
l <- ggplot(data,aes(fill = murder)) #fill color is by murder number for each state. 
l + geom_map(aes(map_id=state), map=map)+ #gives an example of how to use ggplot to create unique maps
  expand_limits(x=map$long, y=map$lat)
```



