# Data Visualization


## Data Visualization in Base R

Basic R plots in base R such as hist() and plot()

```{r}
library(tidyverse)
```

```{r pressure, echo=FALSE}
MN_Data <- read_csv("MN_county_tract_pooled.csv")
plot(pressure)
attach(MN_Data)
hist(median_rent2016)
plot(popdensity2010, nonwhite_share2010)
```

## ggplot2 Data Visualization

the package ggplot has lots of extra plot code to create unique fun plots

```{r}
library(ggplot2)
```

ggplot2 is based on the **grammar of graphics**, it is the idea that you can build graphs from 1) a data set, 2) a set of geoms, 3) and a coordinate system. 

**geoms** are visual marks that represent data points

To display data values, map variables in the data set to **aesthetic** properties of the geom like **size, color, and x and y locations**.


Basic format example: ggplot(data=mpg, aes(x=cty, y=hwy))
"Geoms" represent data points, the geom's aesthetic properties represent variables. 

**ONE VARIABLE CONTINUOUS:**

```{r}
a <- ggplot(mpg, aes(hwy))
a + geom_area(stat = "bin") 
 # [x, y, alpha, color, fill, size]
a + geom_dotplot()
 # [x,y,alpha, color, fill]
a + geom_histogram(binwidth=5)
 # [x,y,alpha, color,fill,linetype,size, weight]
```

  
**ONE VARIABLE DISCRETE:**

```{r}
b <- ggplot(mpg,aes(fl))
b + geom_bar()
  #[x,alpha, color, fill, linetype, size, weight]
```

**TWO VARIABLE CONTINUOUS X, CONTINUOUS Y**

```{r}
f <- ggplot(mpg, aes(cty,hwy))
f + geom_blank() #Blank graph
f + geom_jitter()
 # [x,y,alpha, color, fill, shape, size]
f + geom_point()
 # [x,y,alpha,color,fill, shape, size]
f + geom_smooth(model = lm)
 # [x,y,alpha, color, fill, ,linetype, size, weight]
```
  
**TWO VARIABLE DISCRETE X, CONTINUOUSE Y:**
```{r}
g <- ggplot(mpg,aes(class,hwy))
g + geom_bar(stat = "identity")
  #[x,y,alpha, color, fill, linetype, size, weight]
g + geom_boxplot()
  #[lower, middle, upper, x, ymax, ymin, alpha, color, fill, linetyoe, shape, size, weight]
```
  
**TWO VARIABLE DISCRETE X, DISCRETE Y:**
```{r}
h <- ggplot(diamonds, aes(cut,color))
h + geom_jitter()
  #[x,y,alpha, color, fill, shape, size]
```
  
**TWO VARIABLE CONTINUOUS FUNCTION:**

```{r}
j <- ggplot(economics, aes(date, unemploy))
j + geom_area()
 # [x,y,alpha,color,fill,linetype, size]
j + geom_line()
  #[x,y,alpha,color,linetype,size]
  
```

**MAPS**
```{r}
#install *maps* package
#library(maps)
#data <- data.frame(murder = USArrests$Murder,state=tolower(rownames(USArrests)))
#map <- map_data("state")
#l <- ggplot(data,aes(fill = murder))
#l + geom_map(aes(map_id=state), map=map)+
#  expand_limits(x=map$long, y=map$lat)
 # [map_id, alpha, color, fill, linetype, size]
```

**THEME OPTIONS:**
```{r}
r <- ggplot(mpg, aes(hwy))
r + theme_bw()
r + theme_grey()
r + theme_classic()
r + theme_minimal()
```

**LABELS:**
```{r}

t <- ggplot(mpg, aes(hwy))

t + ggtitle("New Plot Title") #adds a main title above the plot

t + xlab("New X Label") #Change the label on the X axis

t + ylab("New Y Label") #Change the label on the Y axis

t + labs(title="New title", x="New X", y="New Y") #all of the above labels
```


```{r}
ggplot(mpg, aes(hwy, cty)) # This creates a basic graph that you can add layers to to finish it. 

ggplot(mpg, aes(hwy, cty)) +
  geom_point(aes(color = cyl)) +
  geom_smooth(method="lm") +
  coord_cartesian() +
  scale_color_gradient() +
  theme_bw()
```




